Sequential rank agreement methods for comparison of ranked lists

### 2017-03-31 latex/superranker-manuscript.tex
Slowpoke simulation draft done

The manuscript is organized as follows: In the next section we define
sequential rank agreement for multiple ranked lists and discuss how to
handle incomplete lists. In section 3 we present and
discuss approaches to evaluate the results obtained from sequential
rank agreement. Finally we apply the sequential rank agreement to two
Danish ovarian cancer studies before we discuss the findings along
with possible extensions. The approaches presented in this manuscript
are available in the R package \texttt{SuperRanker} which can be found
on CRAN.
% chunk 1

@@

>>>>

The manuscript is organized as follows: In the next section we define
sequential rank agreement for multiple ranked lists and discuss how to
handle incomplete lists. In section 3 we present and discuss
approaches to evaluate the results obtained from sequential rank
agreement. Finally we apply the sequential rank agreement to two
Danish ovarian cancer studies and compare our method to the method of
\citet{hall:schi:2012} in a small sample simulation study before we
discuss the findings along with possible extensions. The approaches
presented in this manuscript are available in the R package
\texttt{SuperRanker} which can be found on CRAN.
% chunk 1

@@

\section{Simulation study --- comparison of list agreements}

We conducted a simulation study to investigate the small sample list
depth agreement properties for our proposed method and to compare with
the method propsed by \citet{hall:schi:2012}. For each of the
simulations we assumed an set of 1000 predictors and a total of 400
observations. The predictors were independent and drawn from a
standard Gaussian distribution. The outcome was also drawn from a
Gaussian distribution with variance 1 but with a mean that depends
additively on the first 15 predictors, i.e.,

$$y_i = \sum_{j=1}^{15}x_{ij} + \epsilon_i,$$

where $y_i$ is the $i$th response, and $x_{ij}$ is the $j$th predictor
for the $i$th measurement, and $\epsilon_i\sim N(0, 1)$. We produce
$L$ ranked lists from a simulated dataset by first generating
non-parametric bootstrap copies of the data, and then for each of the
$L$ bootstrap sample ranking the 1000 predictors according to the size
of their marginal $t$ test statistics. Thus, the simulation results
compare lists that are analysed with the same technique on slightly
different versions of the same data.

In the simulations we focused on two main cases:
\begin{enumerate}
\item The choice of threshold, $q$. Throughout the simulations we
  assume a constant, prespecified threshold function, $q_L$, and try
  different values for the constant. The approach by
  \citet{hall:schi:2012} uses a window size parameter, $K$, to
  represent the maximum difference in rankings before they consider an
  item in two lists to be in agreement.  Because the standard
  deviation --- which is used in the sequential rank agreement
  definition of $\hat{A}_L$--- squares its differences, it tends to
  give more weight to larger rank differences and less weight to
  smaller differences compared to the mean absolute difference.  To
  ease comparison between the sequential rank agreement and the
  \citet{hall:schi:2012} approach we would like to use the same
  threshold for both methods, and in order to do this we use the
  median absolute deviance throughout the simulations in lieu of the
  pooled standard deviation.
 \item The impact of the number of available lists, $L$. Asymptotic
   results are available for both methods as the number of lists
   increases, so this will investigate how much the results might
   change as $L$ increases. For a fixed constant threshold of $q_L=5$
   we increase the number of available lists to investigate the
   sensitivity of the results to $L$.
 \end{enumerate}

 For all combinations we used 1000 simulations to see the distribution
 of the estimated depth of agreement and we also computed the average
 number of different predictors found for the given depth.

 The results for varying threshold is shown in the left-hand plot of
 Figure~\ref{fig:example99}, where the threshold/window size is varied
 from 3 to 10 for a fixed number of lists $L=8$. The double histograms
 show the distribution of the estimated depth of agreement, while the
 bold numbers show the average number of unique predictors found given
 the estimated depth of agreement. We see that the sequential rank
 agrement method results in a substantially lower depth of agreement
 than the \citet{hall:schi:2012} method (denoted topK) and also that
 the average number of unique predictors is markedly smaller. Even
 larger differences are found if we used the Euclidian distance
 instead of the median absolute distance for the sequential rank
 agreement (output not shown).  Not surprisingly, the depth of
 agreement increases when the acceptable threshold for
 agreement/window increases.

 \begin{figure}[htbp]
   \begin{center}
 \includegraphics[width=.49\textwidth]{Ksim.pdf}
 \includegraphics[width=.49\textwidth]{Lsim}
 \end{center}
 \caption{Left panels: Simulation study showing distribution of
   estimated rank agreements for sequential rank agreement and the
   topK for varying thresholds and fixed number of lists $L=8$. The bold numbers are the average
   number of unique predictors in the set found to be in agreement.  Right
   panels: Simulation results for varying number of lists, and with
   fixed threshold of $q=5$.}
 \label{fig:example99}
 \end{figure}

 Figure~\ref{fig:example99} also shows the results when the threshold
 is fixed at $q=5$ and the number of lists increases. The difference
 in agreement between the two methods is still noticeable with the
 sequential rank agreement being more conservative than topK which
 consequently results in fewer predictors among the part of the lists,
 where they are in agreement.  However, the number of lists, $L$,
 seems to have little impact on the distribution of the depth of
 agreement and consequently on the number of different predictors
 among the parts of the lists in agreement.

 For fixed $q$, the two methods are clearly different in the depth
 they find to be in agreement with the sequential rank agreement
 method being the more conservative.
